---
title: "download-all-aihw-data"
author: "Beth Firipis"
date: "2025-07-25"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{download-all-aihw-data}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE # Set to FALSE to prevent long downloads during package build
)
```

```{r, message=FALSE, warning=FALSE}
library(readaihw)
library(dplyr)
```

## Overview

The readaihw package includes functionality to download all available AIHW MyHospitals data in one go, with intelligent caching to avoid unnecessary re-downloads.

## Basic Usage

```{r}
# Download all data to default file (aihw_data.rds)
all_data <- download_all_aihw_data()

# Or specify a custom file location
all_data <- download_all_aihw_data("my_aihw_data.rds")
```

The downloaded data includes: datasets, measures, reporting units, measure categories, hospital mappings, flat data, download timestamp, and version hash for cache validation.

## Caching System

The package includes intelligent caching that detects data updates and avoids redundant downloads:

```{r}
# First download - will fetch fresh data
data1 <- download_all_aihw_data("cached_data.rds")
#> Downloading AIHW data...
#> Saved 6527 datasets, 33 measures, 1423 reporting units

# Second call - uses cached data if current
data2 <- download_all_aihw_data("cached_data.rds")
#> cached_data.rds already exists and is the latest release (downloaded: 2024-01-15 10:30), use download_all_aihw_data(force=TRUE) to force download

# Force fresh download
data3 <- download_all_aihw_data("cached_data.rds", force = TRUE)
#> Downloading AIHW data...
```

## Working with Downloaded Data

```{r}
# Access specific components
head(all_data$datasets)
head(all_data$flat_data)

# Find emergency department data
ed_data <- all_data$flat_data |>
  filter(str_detect(measure_name, "Emergency"))

# Find mental health datasets
mental_health_data <- all_data$datasets |>
  filter(str_detect(tolower(reported_measure_name), "mental"))
```
